Transfer Learning in Visual and Relational Reasoning

### 2020-02-08 Transfer_Learning/cvpr_2nd_submission/cvpr_paper_main.tex
Experiments: basics, feature transfer
Signed-off-by: ibm <tkornuta.ibm@gmail.com>
reasoning transfer moved to a separate file
Signed-off-by: ibm <tkornuta.ibm@gmail.com>
reasoning transfer - COG
Signed-off-by: ibm <tkornuta.ibm@gmail.com>

\usepackage{cvpr}
\usepackage{times}
%\usepackage{epsfig}
\usepackage{graphicx}
\usepackage[font=small]{subcaption}
\captionsetup[figure]{skip=5pt}
\captionsetup[table]{skip=5pt}
\setlength{\belowcaptionskip}{-15pt}  % reduce padding below figure
\usepackage{amsmath, amssymb, amsfonts,bm}
\usepackage{amsthm}

@@

\author{%
	T.S. Jayram\thanks{IBM Research AI.} \and
	Vincent Marois\footnotemark[2] \and
	Tomasz Kornuta\thanks{Nvidia Research.} \and
	Vincent Albouy\thanks{} \and
	Emre Sevgen\thanks{Citrine Informatics.}  \and
	Ahmet Ozcan\footnotemark[2]
}
%
%\author{T.S. Jayram \and  Vincent Marois \and Vincent Albouy \and Tomasz Kornuta  \and Emre Sevgen \and Ahmet Ozcan\\
%IBM Research AI\\
%Almaden Research Center, San Jose, CA 95120, USA\\
%Thomas J. Watson Research Center, Yorktown Heights, NY 10598, USA\\
%{\tt\small {jayram,  asozcan}@us.ibm.com}
%{\tt\small {tkornuta, sesevgen}@gmail.com}
%{\tt\small vincent.marois@ibm.com}
%}

@@

\input{experiments}

>>>>

\usepackage{cvpr}
\usepackage{times}
%\usepackage{epsfig}
\usepackage{graphicx}
\usepackage[font=small]{subcaption}
\captionsetup[figure]{skip=5pt}
\captionsetup[table]{skip=5pt}
%\setlength{\belowcaptionskip}{-15pt}  % reduce padding below figure
\usepackage{amsmath, amssymb, amsfonts,bm}
\usepackage{amsthm}

@@

\author{%
	T.S. Jayram\thanks{IBM Research AI.} \and
	Vincent Marois\footnotemark[2] \and
	Tomasz Kornuta\thanks{NVIDIA.} \and
	Vincent Albouy\thanks{} \and
	Emre Sevgen\thanks{Citrine Informatics.}  \and
	Ahmet Ozcan\footnotemark[2]
}
%
%\author{T.S. Jayram \and  Vincent Marois \and Vincent Albouy \and Tomasz Kornuta  \and Emre Sevgen \and Ahmet Ozcan\\
%IBM Research AI\\
%Almaden Research Center, San Jose, CA 95120, USA\\
%Thomas J. Watson Research Center, Yorktown Heights, NY 10598, USA\\
%{\tt\small {jayram,  asozcan}@us.ibm.com}
%{\tt\small {tkornuta, sesevgen}@gmail.com}
%{\tt\small vincent.marois@ibm.com}
%}

@@

\input{experiments}
\input{experiments_reasoning_transfer}

### 2020-02-08 Transfer_Learning/cvpr_2nd_submission/experiments_reasoning_transfer.tex
reasoning transfer - COG
Signed-off-by: ibm <tkornuta.ibm@gmail.com>

\subsection{Reasoning transfer in CLEVR-CoGenT}
\label{sec:reasoning-clevr}
In the following we have relied on the CoGent-A variant only.
Using the question groups (categories) defined by the authors, we reorganized training and testing splits with the goal of measuring whether mastering reasoning for one group can help with learning of the other.

@@

First we have trained a single SAMNet model and measured its accuracy on each of the task group $t$ separatelly.
Those results, indicated in \cref{fig:cogent_reasoning_transfer} as "Train on All", form baselines for the next two sets of experiments.

Next, we trained and tested 5 SAMNet instances, each on a single task group $t$.
Interestingly, we observe a significant accuracy drop in all task groups aside of the \textit{Query Attribute}.
We hypotesize this results from the fact that mastering visual grounding and recognizing object attributes (being the core of the latter task) is a necessary skill the system needed by the other tasks.
Thus training on \textit{Query Attribute} helps others, whereas without any training on that particular group the system struggles with developing such a skill.

@@

\subsection{Reasoning transfer in COG}
\label{sec:reasoning-cog}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\columnwidth]{../img/plots/cog_reasoning_transfer.pdf}
	\caption{Accuracies of SAMNet when testing its reasoning transfer capabilities on the Canonical variant of COG.}
	\label{fig:COG-reasoning-results}
\end{figure}\vspace{2pt}

Reusing the task hierarchy in \cref{fig:task-groups}, we conduct the following experiments using the Canonical variant of COG to study whether reasoning transfer learning was effective in leveraging information gained by training a task family at a higher level
of the hierarchy:
 \begin{itemize}
 	\compresslist
	\item Train and test SAMNet on each of the 5 task groups at the lowest level of the hierarchy (Single-task training);
	\item Jointly train on:
	\begin{itemize}
		\item \textbf{Group A} and test on each task from the lowest groups (i.e. \textbf{Basic}, \textbf{Obj-Attr} and \textbf{Compare}) separately;
		\item \textbf{Group B} and test separately on \textbf{Spatial} and \textbf{Cognitive};
	\end{itemize}
	\item As a baseline, we compared the above results to the earlier experiment shown in~\cref{fig:samnet_cog_detailed}, which can be viewed as training jointly on \textbf{All} and testing on each group at the leaf level separately (All-task training).
\end{itemize}

The results of these experiments are shown in~\cref{fig:COG-reasoning-results}.

>>>>

\subsection{Reasoning transfer in CLEVR-CoGenT}
\label{sec:reasoning-transfer-clevr}
In the following we have relied on the CoGent-A variant only.
Using the question groups (categories) defined by the authors, we reorganized training and testing splits with the goal of measuring whether mastering reasoning for one group can help with learning of the other.

@@

First we have trained a single SAMNet instance and measured its accuracy on each of the task group $t$ separatelly.
Those results, indicated in \cref{fig:cogent_reasoning_transfer} as "Train on All", form baselines for the next two sets of experiments.

Next, we trained and tested 5 SAMNet instances, each on a single task group $t$.
Interestingly, we observe significant accuracy drops (from 18 up to almost 35 points) in all task groups aside of the \textit{Query Attribute}.
We hypotesize this results from the fact that mastering visual grounding and recognizing object attributes (being the core of the latter task) is a necessary skill the system needed by the other tasks.
Thus training on \textit{Query Attribute} helps others, whereas without any training on that particular group the system struggles with developing such a skill.

@@

\subsection{Reasoning transfer in COG Canonical}
\label{sec:reasoning-transfer-cog}

We conduct the following experiments using the Canonical variant of COG to study whether reasoning transfer learning was effective in leveraging information gained by training a task family at a higher level of the hierarchy.
Order of the performed experiments is analogical to the one from previous chapter.
First column (\cref{fig:cog_reasoning_transfer}, "Train on All") is our baseline, i.e.
we train a single SAMNet instance on all tasks and test on each of the five task groups from the task hierarchy presented in \cref{fig:task-groups}.
Please note that those results are in fact weighted averages when aggregating accuracies of our model on the Canonical variant (\cref{fig:samnet_cog_detailed}).

Next

 \begin{itemize}
 	\compresslist
	\item Train and test SAMNet on each of the 5 task groups at the lowest level of the hierarchy (Single-task training);
	\item Jointly train on:
	\begin{itemize}
		\item \textbf{Group A} and test on each task from the lowest groups (i.e. \textbf{Basic}, \textbf{Obj-Attr} and \textbf{Compare}) separately;
		\item \textbf{Group B} and test separately on \textbf{Spatial} and \textbf{Cognitive};
	\end{itemize}
	\item As a baseline, we compared the above results to the earlier experiment shown in~\cref{fig:samnet_cog_detailed}, which can be viewed as training jointly on \textbf{All} and testing on each group at the leaf level separately (All-task training).
\end{itemize}

The results of these experiments are shown in~\cref{fig:cog_reasoning_transfer}.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\columnwidth]{../img/plots/cog_reasoning_transfer.pdf}
	\caption{Accuracies of SAMNet when testing its reasoning transfer capabilities on the Canonical variant of COG.}
	\label{fig:cog_reasoning_transfer}
\end{figure}
